---
title: "Lesson 5a. Presenting Output Data Analysis"
author: "Assoc. Prof. D. Phillips and N. Uhan"
date: "SA421 -- Simulation Modeling -- Fall 2017"
output: 
  html_notebook: 
    css: css/sa421.css
---

# Setting up your R Markdown file

* You can start with a fresh R Markdown template by selecting __File > New
  File > R Notebook__ in RStudio, or you can modify an existing Rmd file

* Use this as a starting point for your header:

```
---
title: Insert Descriptive Title Here
author: Your Name
date: SA421 -- Simulation Modeling -- Fall 2017
output: html_notebook
---
```

* The last line will ensure that RStudio gives you the option to preview your
  work

# Running your JaamSim file from R

* To run your JaamSim file directly from your R Markdown file, you can use 
  the `system()` function, like this:

```
system('java -jar "Location of JaamSim .jar file" "Location of JaamSim .cfg file" -b')
```

* The `-b` flag tells JaamSim to run the file immediately, without the
  graphical interface

* `05.cfg` in this folder contains the same JaamSim file we used in Lesson 5

* To run this file directly from this R Markdown file, I would write (on my computer):

```{r}
system('java -jar "/Applications/jaamsim/JaamSim2017-09.jar" "05.cfg" -b')
```

* This JaamSim file is configured to write output to `05.dat`. Recall in
  Lesson 5 that we read in this output data as follows:

```{r}
# Read output from simulation
output <- read.table("05.dat", sep="\t", skip=2, 
                     col.names=c("number.of.baristas", "run", "avg.delay"))
```

# Tables with kable

* Like we did in Lesson 5, let's create a data frame containing the summary
  statistics we want:

    - observed sample mean of average delay
    
    - length of 95% confidence intervals

```{r}
# Load dplyr
library(dplyr)

# Group JaamSim output data by number of baristas
# Get sample mean and 95% CI length for each group
# Put into a separate data frame so we can access it later
summary.avg.delay <- output %>%
  group_by(number.of.baristas) %>%
  summarise(sample.mean = mean(avg.delay), ci.length.95 = 2*1.96*sd(avg.delay)/sqrt(n()))

# Show the summary data frame 
summary.avg.delay
```

* This data frame contains the information we want, but is not
  publication-worthy: the column headings should be in plain English, not R
  variable names
    
* We can convert this data frame to a proper table using __kable__, which is a
  part of the knitr library:

```{r}
# Load knitr
library(knitr)

# Convert summary data frame into a proper table
kable(summary.avg.delay,
      col.names=c("Number of baristas", "Sample mean", "95% CI Length"))
```

* Alternatively, we can chain the sample mean and CI computation with
  `kable()` like this:

```{r}
output %>%
  group_by(number.of.baristas) %>%
  summarise(sample.mean = mean(avg.delay), ci.length.95 = 2*1.96*sd(avg.delay)/sqrt(n())) %>%
  kable(col.names=c("Number of baristas", "Sample mean", "95% CI Length"))
```


# Plots with ggplot2

* What if we want to describe these results visually?

* Let's plot the observed sample mean of the average delay and the
  corresponding confidence interval ranges using __ggplot2__

* ggplot2 is a system in R for creating graphics

    - [Homepage](http://ggplot2.tidyverse.org/index.html) and [documentation](http://ggplot2.tidyverse.org/reference/) for ggplot2

* Roughly speaking, building a plot with ggplot2 looks like this:

```
data +
    geometry to represent the data (e.g. point plot? bar graph?)  
        with mapping of data to geometry properties 
        (e.g. x-coordinates, y-coordinates, shapes and colors) + 
    scaling of ranges (e.g. regular? log?) + 
    axis labels and title
```

* For example, a simple point plot of the mean average delay:

```{r}
# Load ggplot2
library(ggplot2)

# An example of a point plot in ggplot2
ggplot(summary.avg.delay) +                       # Which data?
  geom_point(aes(x = number.of.baristas,          # What kind of plot? 
                 y = sample.mean)) +              #   How does the data map to the plot? 
  xlab("Number of baristas") +                    # x-axis label
  ylab("Observed sample mean of average delay") + # y-axis label
  ggtitle("Average Delay at the Barista Queue")   # title label
```

* Maybe this would look better as a bar graph? 

    - `stat = "identity"` tells `geom_bar()` to use the y values as the height
      of the bars

```{r}
# An example of a bar graph in ggplot2
ggplot(summary.avg.delay) +
  geom_bar(aes(x = number.of.baristas, 
               y = sample.mean), 
           stat = "identity") +                   
  xlab("Number of baristas") + 
  ylab("Observed sample mean of average delay") + 
  ggtitle("Average Delay at the Barista Queue")
```

* Or maybe... we should have points showing the sample means, along with error
  bars showing the confidence interval ranges?
    
    - It would also be nice to revise the y-axis scale as well, so some of the 
      differences in values are easier to see

    - `scale_y_sqrt()` stretches the scale on the y-axis at smaller values

    - Instead of `xlab()`, `ylab()` and `ggtitle()`, you can also use `labs()`
      to specify the axis labels and title
    
    - You can also specify a caption and subtitle with `labs()`

```{r}
# An example of a point plot with error bars in ggplot2
ggplot(summary.avg.delay) +
  geom_point(aes(x = number.of.baristas, 
                 y = sample.mean), 
             stat = "identity") + 
  geom_errorbar(aes(x = number.of.baristas, 
                    ymin = sample.mean - ci.length.95/2, 
                    ymax = sample.mean + ci.length.95/2)) +
  scale_y_sqrt() +
  labs(x = "Number of baristas",
       y = "Observed sample mean of average delay",
       title = "Average Delay at the Barista Queue", 
       caption = "The error bars correspond to 95% confidence intervals.")
```

# Some tips on presenting your output data analysis

* When presenting your output data analysis, __your goal is to accurately 
  describe the output data, not make decisions or judgements based on the data__

* Present the numerical values of your output data analysis using tables or
  graphs (or both)

* Give your tables descriptive column headings

* Label the axes of your graphs and give your graphs descriptive titles

* __Accompany your tables and graphs with words:__

    - Introduce each table or graph, e.g.,
    
        > The graph below shows the average delay predicted by our simulation
        > for different numbers of baristas working in the cafe.

    - Describe each table or graph, but __do not__ simply repeat the numerical 
      values in your tables and graphs
    
    - Instead, point your reader to important trends while referring to your
      tables or graphs, e.g.,
    
        > As we can see in the table above, as the number of baristas goes up,
        > the observed sample mean of the average delay goes up as well.
    
    - You may also want to point out the approximate magnitude in changes, if
      applicable, e.g.,
    
        > Each additional barista results in an approximately 200% increase in
        > the predicted average delay.
        
* In the case of the NCB problem, the observed sample mean of the average
  delay is your simulation study's _prediction_ of the mean average delay
      
      - So an appropriate synonym for "observed sample mean of the average 
        delay" would be "predicted mean average delay."
